---
title: "Isoscape assignment SRW - all breeding grounds"
author: "Solene Derville"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
   html_document:
    highlight: pygments
    keep_md: yes
    number_sections: true
    toc: true
    toc_float:
      collapsed: false
---

```{r, echo=F, result="hide", include=F}
lapply(c("plyr", "tidyverse", "raster", "sf", "ncdf4", "rnaturalearth", "rnaturalearthdata", "ggsflabel", "gridExtra", "foreach", "doParallel", "patchwork", "stars"), library, character.only=TRUE)

## proj4string for coordinates systems to use throughout code
proj_lon360 <- "+proj=longlat +ellps=WGS84 +lon_wrap=180 +datum=WGS84 +no_defs"
proj_antarctic <- "+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +k=1 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"

sf::sf_use_s2(FALSE)
# land polygons for plots
world <- ne_countries(scale = "medium", returnclass = "sf") 
world <- st_crop(world, xmin = -180, xmax = 180, ymax = -90, ymin = 0)
world <- st_transform(world, proj_antarctic)
```

Dataset with individual samples collected in breeding grounds
```{r}
srw.data <- read.csv("../Data/SRW_Sample_isotope_analysis_5March21_SD.csv", stringsAsFactors = F) %>% 
  dplyr::select(Year, Month, D13C, D15N, Locality1, Lat, Long, AgeClass) %>% 
  filter(Locality1 != "Auckland Islands", Locality1 != "NZ Mainland") # remove the old NZ data

# new data for NZ - Auckland Islands replaced the old one
nz.data <- read.csv("../Data/NZ_SRW_SI_UniquePerYearNov2021_SD.csv") %>% 
  dplyr::select(Year, Month, D13C, D15N, Locality1, Lat, Long, AgeClass)

srw.data <- rbind(srw.data, nz.data)

## remove calves and data outside the months of July-oct
srw.data <- srw.data %>% 
  filter(!(AgeClass %in% c("calf")) & Month >= 7 & Month <= 10)


# how many samples
nrow(srw.data)

# separate the Argentina data into two different subgroups as well as having it as one complete breeding stock
argentina <- srw.data %>% 
  filter(Locality1 == "Argentina") %>% 
  mutate(Locality1 = case_when(D15N > 10 ~ "Argentina_d15Nhigh",
                               D15N <= 10 ~ "Argentina_d15Nlow"))
table(argentina$Locality1)
srw.data <- rbind(srw.data, argentina)

# add X at the beginning of the sample name because the raster function cannot use numerical variables as names for the layers
# sampleID had too many special characters that were not recognized in the raster function
srw.data$Sample  <- paste("X", seq(1, nrow(srw.data), 1), sep = "")

# verify that Sample numbers are unique to each row in the dataframe
nrow(srw.data)
length(unique(srw.data$Sample))

# convert data to spatial format and project
srw.sf <- st_as_sf(srw.data, coords = c("Long", 'Lat'), crs = 4326) %>%
  st_transform(crs = st_crs(proj_antarctic)) 

# breeding sites names in spatial format
labels_sites <- srw.sf %>%
  group_by(Locality1) %>%
  summarize()
# make the names matches with those in the raster
labels_sites$Locality <-labels_sites$Locality1 
```

Load phytoplankton base rasters
```{r}
load("../2-Tunning_with_tracks/Outputs/D13C_ras.RData")
load("../2-Tunning_with_tracks/Outputs/D15N_ras.RData")
```


Set up a plot template
```{r}
gbase <- ggplot() +
  scale_fill_viridis_c(na.value = "transparent", option = "viridis", 
                       name = "Mean probability of origin \nacross individuals", direction = -1) +
  coord_sf(label_graticule = "") +
  theme_grey() +
  theme(legend.position = "bottom") +
  xlab("") + ylab("")
```

# Creating individual rasters for assignment

## Check out temporal distribution of samples
```{r}
g <- ggplot(srw.data, aes(as.factor(Month), fill = Locality1)) +
  geom_bar() +
  facet_wrap(~Locality1, scales = "free_y") +
  ylab("Number of samples") +
  xlab("Month")
g
ggsave("./Outputs/srw.data_sample_distribution.png", width = 7, height = 5)
```


## Averaging rasters per Individual

For each locality, retrieve relevant monthly layers of D13C and D15N (3 to 5 months prior to date of sample collection). This chunk needs a few minutes to run on my computer.
```{r, eval = F}
# the names of the layer are funky because the time in the netcdf was in decimal years. But we can still easily access the year so that's enough
head(names(D13C_ras))
head(substr(names(D13C_ras), 2, 5))
# For this I assumed that the netcdf contained one layer per month per year since January 1992. If there are gaps in this time series, it will mess up the assignment in the loop below. Important to check this is correct.
srw.ras.nonmasked <- dlply(srw.data, ~Sample, function(d){ # for each sample
  # month of the sample
  m <- d$Month[1]
  # selecting the relevant layers from that year
  n <- names(D13C_ras)[substr(names(D13C_ras), 2, 5) == d$Year[1]]
  # selecting only the fifth, fourth, and third month prior to sample
  n <- n[(m-5):(m-3)]
  # retrieving these layers from the rasterbrick
  ras13mean <- mean(D13C_ras[[n]], na.rm = F)
  ras13sd <- calc(D13C_ras[[n]], function(r) sd(r))
  ras15mean <- mean(D15N_ras[[n]], na.rm = F)
  ras15sd <- calc(D15N_ras[[n]], function(r) sd(r))
  # stack the four rasters together
  r <- stack(ras13mean, ras13sd, ras15mean, ras15sd)
  # set names to something more explicit
  names(r) <- c("D13Cmean", "D13Csd", "D15Nmean", "D15Nsd")
  return(r)
})

# Saving these individual Rasters stack (without masking)
save(srw.ras.nonmasked, file = "./Outputs/srw.ras.nonmasked.RData")
```

Check out an example
```{r, fig.cap = "3-month average of phytoplankton isoscape prior to sample taken in Argentina breeding grounds.", fig.align="center", fig.height = 8, fig.width = 10}
load("./Outputs/srw.ras.nonmasked.RData")
plot(srw.ras.nonmasked[["X188"]], main = "Argentina individual example")
```

```{r, fig.cap = "3-month average of phytoplankton isoscape prior to sample taken in Auckland Islands breeding grounds.", fig.align="center", fig.height = 8, fig.width = 10}
plot(srw.ras.nonmasked[["X630"]], main = "Auckland Islands individual example")
```

## Masking

Average position of samples by breeding stock
```{r}
srw.data.breeding <- srw.data %>% 
  group_by(Locality1) %>% 
  dplyr::summarize(lon = mean(Long, na.rm = T),
            lat = mean(Lat, na.rm = T))
srw.data.breeding
```

Convert breeding region positions to spatial object, project and buffer around with 6500 km. Then crop out the parts of the feeding bubbles that are > 30°S and smaller parts of the bubble that might come up on the other side of the antarctic continent...
```{r}
# create circle polygon centered on pole and up to 30°S latitude
crop_circle <- data.frame(lon = c(0), lat = c(-90)) %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>% 
  st_transform(crs = st_crs(proj_antarctic)) %>% 
  st_buffer(dist = 60*111e3) # one degree of latitude is 111 km

# create buffers around breeding regions
srw.buf <- st_as_sf(srw.data.breeding, coords = c("lon", 'lat'), crs = 4326) %>%
  st_transform(crs = st_crs(proj_antarctic)) %>%
  st_buffer(dist = 6500e3)
 
# crop at 30°S
srw.buf <-  st_intersection(srw.buf, crop_circle)

srw.buf
```

```{r, fig.cap = "Basic 6500 km feeding bubble cropped at 30°S.", fig.align="center", fig.height = 6, fig.width = 10}
ggplot() +
  geom_sf(data = srw.buf, aes(fill = Locality1), alpha = 0.8) +
  geom_sf(data = world, col = "transparent", fill = "grey15") +
  facet_wrap(~Locality1, ncol = 4) +
  coord_sf(label_graticule = "") +
  theme_grey() +
  theme(legend.position = "bottom") +
  xlab("") + ylab("")
```

We also noticed that the phytoplankton isoscapes showed elevated and unusual values in coastal waters around the Antarctic continent (e.g. Ross Sea) so we decided to remove these waters from the analysis.
```{r, warning = F}
## crop waters on continental shelf with a buffer around land
antarctica_buf <- world %>% 
  filter(continent == "Antarctica") %>% # select antarctica continent
  st_cast("POLYGON") # convert multipolygon to polygons
# calculate surface area of polygons
antarctica_buf$area <- st_area(antarctica_buf)
# select the mainland only and buffer
antarctica_buf <- antarctica_buf %>% 
  filter(area == max(antarctica_buf$area)) %>% 
  st_buffer(200e3)
plot(antarctica_buf['continent'])

srw.buf <-  st_difference(srw.buf, antarctica_buf) %>% 
  dplyr::select(Locality1) # remove useless columns


# for argentina and brazil also remove the Pacific side of the bubble
# tracing a line between Navarino island (-55.169070, -67.611678) and Antarctica (-79, -67.611678)
exclusion_pacific <- st_read("../Data/polygon_exclusion_pacific.shp") %>% 
  st_transform(proj_antarctic)

# remove pieces of bubble on the other side of the antarctic continent...
# for brazil remove the pacific side...
srw.buf.list <- list()
for(i in 1:nrow(srw.buf)){
  srw.buf.i <- srw.buf[i, ]
  # if Brazil or Argentina, remove pacific part of the buffer
  if(srw.buf.i$Locality1 %in% c("Brazil", "Argentina", "Argentina_d15Nhigh","Argentina_d15Nhigh", "Argentina_d15Nlow")){
    srw.buf.i <- st_difference(srw.buf.i, exclusion_pacific) 
  }
  srw.buf.i <- srw.buf.i %>% 
    st_cast("POLYGON", warn = F) # convert multipolygon to polygons
  # calculate surface area of polygons
  srw.buf.i$area <- st_area(srw.buf.i)
  # select the main part of the bubble
  srw.buf.i <- srw.buf.i %>% 
    filter(area == max(srw.buf.i$area))
  srw.buf.list <- c(srw.buf.list, list(srw.buf.i))
}
srw.buf <- ldply(srw.buf.list, rbind) %>% 
  st_as_sf(crs = proj_antarctic)

save(srw.buf, file = "./Outputs/srw.buf.RData")
```

```{r, fig.cap = "Feeding bubble outside which isoscape assignment will be masked.", fig.align="center", fig.height = 6, fig.width = 10}
ggplot() +
  geom_sf(data = srw.buf, aes(fill = Locality1), alpha = 0.8) +
  geom_sf(data = world, col = "transparent", fill = "grey15") +
  facet_wrap(~Locality1, ncol = 4) +
  coord_sf(label_graticule = "") +
  theme_grey() +
  theme(legend.position = "bottom") +
  xlab("") + ylab("")
```

Apply same buffer to all individual rasters per breeding ground for masking
```{r, eval = T}
srw.ras.masked <- alply(names(srw.ras.nonmasked), .margins = 1, function(n){
  # go retrieve the Locality
  l <- srw.data[srw.data$Sample == n, "Locality1"]
  # rotate individual raster will just convert back -180_180 lon
  ind <- rotate(srw.ras.nonmasked[[n]])
  # now project to the antarctic centered crs
  ind <- projectRaster(ind, crs = crs(proj_antarctic))
  # apply mask to set cells outside the buffers to NA
  masked_ras <- raster::mask(ind, st_as_sf(srw.buf[srw.buf$Locality1 == l, ]))
  return(masked_ras)
})
names(srw.ras.masked) <- names(srw.ras.nonmasked)

# Saving these individual Rasters stack
save(srw.ras.masked, file = "./Outputs/srw.ras.masked.RData")
```

Check out an example
```{r, fig.width = 10}
load("./Outputs/srw.ras.masked.RData")
plot(srw.ras.masked[["X1000"]]) # auckland islands
plot(srw.ras.masked[["X5"]]) # south africa
plot(srw.ras.masked[["X203"]]) # argentina
```

# Assignment function by individual

## TDF settings

I am applying the best TDF we found in the Auckland Islands - Argentina satellite tracking analysis: TDF is 4.5 for d15N and 3 for d13C
```{r, eval = F}
srw.ras.masked.corrected <- llply(srw.ras.masked, function(r){
    r$D13Cmean <- r$D13Cmean + 3
    r$D15Nmean <- r$D15Nmean + 4.8
    return(r)
})
save(srw.ras.masked.corrected, file = "./Outputs/srw.ras.masked.corrected.RData")
```

## Constant parameters

```{r}
load("./Outputs/srw.ras.masked.corrected.RData")
## SD_C
## within individual error: δ13C: 0.45±0.57‰ estimated by Carroll et al in prep
SD_C<-0.45
      
## SD_N
## within individual error: δ15N: 0.38±0.42‰ estimated by Carroll et al in prep
SD_N <-0.38

## SD_TEF_C1
## variance due to trophic enrichment factor or rescaling factor: phyto to zooplankton
SD_TEF_C1 <- 0.3

## SD_TEF_N1: phyto to zooplankton
## variance due to trophic enrichment factor or rescaling factor
SD_TEF_N1 <- 0.3

## SD_TEF_C2: zooplankton to whale
## variance due to trophic enrichment factor or rescaling factor
SD_TEF_C2 <- 0.38

## SD_TEF_N2: zooplankton to whale
## variance due to trophic enrichment factor or rescaling factor
SD_TEF_N2 <- 0.30


## rho
## correlation between C and N, calculated on a global scale 
srw.data_rho <- srw.data %>% 
  filter(!(Locality1 %in% c("Argentina_d15Nhigh", "Argentina_d15Nlow")))
nrow(srw.data_rho)
vcm<-cov(cbind(srw.data_rho$D13C, srw.data_rho$D15N))
VarCovar <- vcm
rho <- VarCovar[1,2]/(sqrt(VarCovar[1,1])*sqrt(VarCovar[2,2]))
rho
```

## Assignment model

From Hannah's code, this is the assignment part. I created a function from it, called Fun_assign
```{r}
Fun_assign <- function(individual.isotopes, ras.isoscapes){
  for (i in 1:nrow(individual.isotopes)){# for each individual in your file
    # id of that individual
    name_n <- as.character(individual.isotopes$Sample[i])
    # select the right rasters for that individual
    ras <- ras.isoscapes[[name_n]]
    # select the isotope values for that individual
    x <- individual.isotopes$D13C[i]
    y <- individual.isotopes$D15N[i]
    ## in Hannah's code the next two lines aren't calculated per individual, but if we are generating the raster and its error each loop it needs to be calculated
    ## each time. Two of these parameters are fixed - within individual and TEF variance, but the SD raster will change
    error_x <- sqrt((ras$D13Csd)^2 + (SD_C)^2 + (SD_TEF_C1)^2+ (SD_TEF_C2)^2)
    error_y <- sqrt((ras$D15Nsd)^2 + (SD_N)^2 + (SD_TEF_N1)^2+ (SD_TEF_N2)^2)
    ## assignment code: added in year here too, but you can tweak it how you want to call that variable however you like!
    assign <- (1/(2*pi*error_x*error_y*sqrt(1-rho^2))) * (exp(-1/(2*(1-rho^2)) * ((((x-ras$D13Cmean)^2)/(error_x^2)) + (((y-ras$D15Nmean)^2)/(error_y^2))                                                                              -((2*rho*(x-ras$D13Cmean)*(y-ras$D15Nmean) / (error_x*error_y)))))) 
   writeRaster(assign, 
      filename = paste0("./Outputs/assignments/assign_", name_n, ".grd"), bylayer=F, overwrite=T)
  }
}

# run the function over all samples!
# outputs are saved directly to disk in the a folder > Outputs > assignments
# saved as grd files
Fun_assign(individual.isotopes = srw.data, ras.isoscapes = srw.ras.masked.corrected)
```


# Breeding stock analysis

## Averaging at breeding ground scale

Rescaling individual rasters within a breeding ground so that the sum of probabilities within the bubble is 1. Then averaging individuals together per breeding grounds.
```{r, eval = F}
d_ply(srw.data, ~Locality1, function(d){
      # select names of individuals of interest
      ind <- as.character(d$Sample)
      # for each individual and correction setting, go get raster
      files <- paste0("./Outputs/assignments/assign_", ind, ".grd")
      ras <- alply(files, .margins = 1, raster)
      ras <- stack(ras)
      names(ras) <- ind
      # rescale the map to sum to 1
      for (i in ind){
        ras[[i]] <- ras[[i]]/sum(values(ras[[i]]), na.rm = T)
      }
      # average across all individuals sampled in the same breeding ground
      ras_mean <- mean(ras, na.rm = T)
      # rename the layer with name of the locality = breeding ground
      names(ras_mean) <- d$Locality1[1]
      # save to disk as grd file
      raster::writeRaster(ras_mean, 
        filename = paste0("./Outputs/breeding_grounds/assign_", d$Locality1[1], ".grd"), 
        bylayer=F, overwrite=T)
})

```

Read the breeding ground raster averages and save to a raster stack
```{r}
# vector of unique locality names
locs <- unique(srw.data$Locality1)
# list files
files <- paste0("./Outputs/breeding_grounds/assign_", locs,".grd")
# read rasters to list
ras <- alply(files, .margins = 1, raster)
# stack together
breeding.ras <- stack(ras)
names(breeding.ras) <- locs

save(breeding.ras, file = "./Outputs/breeding.ras.RData")
```

Convert to dataframe
```{r}
breeding.df <- as.data.frame(breeding.ras)
breeding.df[c("x", "y")] <- coordinates(breeding.ras)
breeding.df <- breeding.df %>% 
  pivot_longer(cols = c(South.Africa, Brazil, Argentina, Argentina_d15Nlow, Argentina_d15Nhigh, Southwest.Australia,  Southeast.Australia, Auckland.Islands, NZ.Mainland), names_to = 'Locality', values_to = 'proba_origin') %>% 
    filter(!is.na(proba_origin))

save(breeding.df, file = "./Outputs/breeding.df.RData")
```

```{r, fig.cap = "Maps of mean assignments by breeding stock", fig.align="center", fig.height = 8, fig.width = 10, warning = F, message = F}
g <- gbase +
  geom_tile(data = breeding.df, aes(x, y, fill = proba_origin)) +
  geom_sf(data = world, col = "transparent", fill = "grey15") +
  facet_wrap(~Locality, ncol = 4)
g
```

## Classify P(origin) into binomial variable

```{r, fig.cap = "Distribution of the probability of origin across each locality's feeding bubble. Note: Probabilities were rescaled to sum to one for each individual assignment map, before being averaged across individuals of the same breeding stock", fig.align="center", fig.height = 8, fig.width = 10, warning = F, message = F}

# calculate 75 % quantile per locality
thresh <- breeding.df %>% 
  group_by(Locality) %>% 
  dplyr::summarize(proba_threshold =
                    round(quantile(proba_origin, 0.75), 6))

# plot histogram with 7(% quantile)
ggplot(breeding.df, aes(proba_origin)) +
  geom_histogram() +
  facet_wrap(~Locality, scales = "free_x") +
  geom_vline(data = thresh, 
             aes(xintercept =proba_threshold)) +
  geom_text(data = thresh, aes(
    x = proba_threshold + 0.022, 
    y = 400, 
    label = paste0("Threshold 75% = ",proba_threshold)), 
    size = 3)
```

Use threshold to classify P(origin) into a binomial variable using two thresholds: the best 25 % is the core of the distribution whereas the best 50 % is a wider estimate of the feeding zone
```{r}
# bin for each breeding raster
breeding.bin <- breeding.df %>% 
  group_by(Locality) %>% 
  mutate(feeding0.25 = ifelse(proba_origin >= quantile(proba_origin, 0.75, na.rm = T), 1, NA),
         feeding0.50 = ifelse(proba_origin >= quantile(proba_origin, 0.5, na.rm = T), 1, NA)) %>% 
  filter(!(is.na(feeding0.25) & is.na(feeding0.50)))

save(breeding.bin, file = "./Outputs/breeding.bin.RData")
```

```{r, fig.cap = "Population level feeding areas", fig.align="center", fig.height = 8, fig.width = 8, warning = F}
g <- ggplot() +
  coord_sf(label_graticule = "") +
  theme_light() +
  theme(legend.position = "") +
  xlab("") + ylab("") +
  geom_tile(data = breeding.bin[!is.na(breeding.bin$feeding0.50), ], aes(x, y, fill = Locality), alpha = 0.5) +
  geom_tile(data = breeding.bin[!is.na(breeding.bin$feeding0.25), ], aes(x, y, fill = Locality), alpha = 1) +
  geom_sf(data = world, col = "transparent", fill = "grey15") +
  facet_wrap(~Locality, ncol = 3)

ggsave(g, file = "./Outputs/Population_feeding_grounds.png", width = 8, height = 8)
g
```

```{r, fig.cap = "Population level feeding areas", fig.align="center", fig.height = 3, fig.width = 7, warning = F}
g <- ggplot() +
  coord_sf(label_graticule = "") +
  theme_light() +
  theme(legend.position = "right") +
  scale_color_discrete(guide = "none") +
  xlab("") + ylab("") +
  geom_tile(data = breeding.bin, aes(x, y, fill = Locality), alpha = 0.7) +
  geom_sf(data = world, col = "transparent", fill = "grey15") +
  geom_sf_label_repel(data = labels_sites, 
         aes(label = Locality, col = Locality), nudge_x = -10e5, nudge_y = 10e3, segment.size = 0, box.padding = 1, size = 2, force = 1)
g
ggsave(g, file = "./Outputs/Population_feeding_grounds_overlap.png", width = 7, height = 5)
```

## Calculate overlap across breeding stocks

Overlap is calculated across the general feeding areas (50 % contours) assigned to each breeding stock
```{r}
# convert the binomial assignments back to a spatial point format
allfeeding <- breeding.bin %>% 
      filter(!is.na(feeding0.50)) %>% 
      dplyr::select(-c(proba_origin, feeding0.25, feeding0.50)) %>% 
      st_as_sf(coords = c("x", "y"), crs = st_crs(proj_antarctic)) 
# st_agr just prevents warning in st_intersection
st_agr(allfeeding) = "constant"

  
breeding.overlap <- breeding.bin %>% 
  group_by(Locality) %>% 
  nest() %>% 
  mutate(data = map(data, function(d){ #select the feeding zones 50 % in each locality
    d %>% 
      filter(!is.na(feeding0.50)) %>% 
      dplyr::select(-c(proba_origin, feeding0.25, feeding0.50)) %>% 
      st_as_sf(coords = c("x", "y"), crs = st_crs(proj_antarctic))
  }),
  overlap_feeding = map(data, function(d){ #intersect the full dataset of 50 % feeding ground with each feeding locality group of cells
    st_agr(d) = "constant"
    d %>% 
      st_intersection(allfeeding, warning = F)
  }),
  percent_overlap = map2(data, overlap_feeding, function(x, y){
    y %>% 
      group_by(Locality) %>% 
      summarize(nb_overlap_cells = n(), # nb of overlap cells by locality
                percent_overlap = round(nb_overlap_cells / nrow(x) *100, 0)) %>% # divide by total number of cells
      rename(Locality_overlapping = Locality)
  }))

# set order of locality in which we want them to appear in the csv
locs <- c("Argentina", "Argentina_d15Nlow", "Argentina_d15Nhigh", "Brazil", "South.Africa", "Southwest.Australia", "Southeast.Australia", "NZ.Mainland", "Auckland.Islands")

# pull the relevant data from the tibble and pivot to create a table to export as csv
breeding.overlap %>% 
  unnest(percent_overlap) %>% 
  dplyr::select(-c(data, overlap_feeding, nb_overlap_cells, geometry)) %>% 
  mutate(Locality = factor(Locality, levels = locs)) %>%
  arrange(Locality) %>% # order the rows
  pivot_wider(names_from = Locality_overlapping, values_from = percent_overlap) %>%
  dplyr::select(locs) %>% # reorder the columns in same order as rows
  replace(is.na(.), 0) %>% # NA means there was zero overlap between the two breeding stocks
  print(n = Inf) %>% 
  write.csv(file = "./Outputs/Table_overlap_breedingstocks.csv", row.names = F)
```

## Individual level maps

Now instead of averaging individual maps by breeding stocks, I will just load them, rescale them to sum to 1 and return them in a list called ind_assign
```{r}
ind_assign <- dlply(srw.data, ~Locality1, function(d){
  # select names of individuals of interest
  ind <- as.character(d$Sample)
  # for each individual go get raster
  files <- paste0("./Outputs/assignments/assign_", ind,".grd")
  ras <- alply(files, .margins = 1, raster)
  ras <- stack(ras)
  names(ras) <- ind
  # rescale the map to sum to 1
  for (i in ind){
    ras[[i]] <- ras[[i]]/sum(values(ras[[i]]), na.rm = T)
  }
  return(ras)
})
```

Bin all individual rasters at 50% quantile, sum and calculate percent of individuals assigned to each pixel
```{r}
## bin the individual rasters and sum them by locality1
ind_bin_sum_df <- ldply(ind_assign, .id = "Locality1", function(r){
  for (i in names(r)){
    q <-  quantile(values(r[[i]]), 0.5, na.rm = T)
    reclass <- matrix(c(0, q, 0,
                        q, Inf, 1),
                        ncol = 3,
                        byrow = TRUE)
    r[[i]] <- reclassify(r[[i]], reclass)
  }
  # sum individual layers after binned
  r_sum <- sum(r, na.rm = T)
  # convert to dataframe
  df <- as.data.frame(r_sum)
  df[c("x", "y")] <- coordinates(r_sum)
  df <- df %>% 
    rename(sum_ind_feeding = layer) %>% 
    filter(!is.na(sum_ind_feeding), sum_ind_feeding != 0) # remove cells with no assignment
  return(df)
})

## join the dataframe with the number of samples per locality1 so we can convert sum_ind_feeding into percent_ind_feeding
nb_samples <- srw.data %>% 
  group_by(Locality1) %>% 
  dplyr::summarize(nb_ind = length(unique(Sample)))
nb_samples
ind_bin_sum_df <- ind_bin_sum_df %>% 
  join(nb_samples, by = "Locality1") %>% 
  mutate(percent_ind_feeding = round(sum_ind_feeding / nb_ind *100, digits = 0))
# check the distrib of the percentages
summary(ind_bin_sum_df$percent_ind_feeding)

save(ind_bin_sum_df, file = "./Outputs/ind_bin_sum_df.RData")
```

```{r, fig.cap = "Population level feeding areas", fig.align="center", fig.height = 3, fig.width = 7, warning = F}
g <- ggplot() +
  coord_sf(label_graticule = "") +
  theme_grey() +
  theme(legend.position = "bottom") +
  xlab("") + ylab("") +
  geom_tile(data = ind_bin_sum_df, aes(x, y, fill = percent_ind_feeding), alpha = 1) +
  geom_sf(data = world, col = "transparent", fill = "grey15") +
  scale_fill_viridis_c() +
  facet_wrap(~Locality1)
ggsave(g, file = "./Outputs/Individual_feeding_grounds.png", width = 8, height = 8)
g
```


# Decadal analysis

## Sample distribution across decades

```{r}
# create new column to indicate what decade the sample was collected in
srw.data <- srw.data %>% 
  mutate(decade = case_when(Year < 2000 ~ "1994-1999",
                            Year >= 2000 & Year <= 2009 ~ "2000-2009",
                            Year >= 2010 ~ "2010-2020")) 

# plot the temporal distribution of samples
g <- ggplot(srw.data, aes(Locality1, Year, col = decade)) +
    geom_jitter(width = 0.2, height = 0, alpha = 0.6) +
    theme_light() +
    xlab("Breeding population")
ggsave(g, file = "./Outputs/samples_by_decades.png", width = 10, height = 5)
g
```

## Averaging by breeding ground x decade

Rescaling individual rasters to scale to 1. Then averaging individuals together per breeding grounds x decade.
```{r, eval = T}
d_ply(srw.data[!(srw.data$Locality1 %in% c("Argentina", "Argentina_d15Nhigh", "Argentina_d15Nlow")), ], .(Locality1, decade), function(d){
      # select names of individuals of interest
      ind <- as.character(d$Sample)
      # for each individual and correction setting, go get raster
      files <- paste0("./Outputs/assignments/assign_", ind, ".grd")
      ras <- alply(files, .margins = 1, raster)
      ras <- stack(ras)
      names(ras) <- ind
      # rescale the map to sum to 1
      for (i in ind){
        ras[[i]] <- ras[[i]]/sum(values(ras[[i]]), na.rm = T)
      }
      # average across all individuals sampled in the same breeding ground
      ras_mean <- mean(ras, na.rm = T)
      names(ras_mean) <- d$Locality1[1]
      raster::writeRaster(ras_mean, 
        filename = paste0("./Outputs/breeding_grounds_decade/assign_", d$Locality1[1],"_", d$decade[1], ".grd"), 
        bylayer=F, overwrite=T)
})

```

Read the breeding ground x decade raster averages and save to a raster stack
```{r}
# retrieve the names of the .grd files by decade by breeding ground
ff <- list.files("./Outputs/breeding_grounds_decade/", pattern = "\\.grd$")
# read and convert to a dataframe
decades.df <- ldply(ff, function(file){
  # read raster
  ras <- raster(paste0("./Outputs/breeding_grounds_decade/", file))
  # convert to dataframe
  df <- as.data.frame(ras)
  names(df) <- "proba_origin"
  df[c("x", "y")] <- coordinates(ras)
  name_file <- strsplit(file, "_")[[1]]
  # retrieve the locality and the decade from the name of the file and add in separate columns to the dataframe
  df$Locality <- name_file[2] 
  df$decade <- strsplit(name_file[3], ".grd")[[1]][1]
  return(df)
})

save(decades.df, file = "./Outputs/decades.df.RData")

```

## Classify P(origin) into binomial variable

```{r}
decades.bin <- decades.df %>% 
  group_by(Locality, decade) %>% 
  mutate(feeding0.25 = ifelse(proba_origin >= quantile(proba_origin, 0.75, na.rm = T), 1, NA),
         feeding0.50 = ifelse(proba_origin >= quantile(proba_origin, 0.5, na.rm = T), 1, NA)) %>% 
  filter(!(is.na(feeding0.25) & is.na(feeding0.50)))

save(decades.bin, file = "./Outputs/decades.bin.RData")
```

```{r, fig.cap = "Population level feeding areas by decades", fig.align="center", fig.height = 9, fig.width = 8, warning = F}
g <- ggplot() +
  coord_sf(label_graticule = "") +
  theme_light() +
  theme(legend.position = "") +
  xlab("") + ylab("") +
  geom_tile(data = decades.bin[!is.na(decades.bin$feeding0.50), ], aes(x, y, fill = Locality), alpha = 0.7) +
  geom_sf(data = world, col = "transparent", fill = "grey15") +
  facet_wrap(Locality~decade, ncol = 4)
g
ggsave(g, file = "./Outputs/Population_decade_feeding_grounds.png", width = 8, height = 9)
```

# Whaling data

## Load Yankee and Soviet whaling data

```{r}
files <- as.list(c("../Data/Historial Whaling data/R/Yankee_Whaling_Data_aowl_SRW.csv", 
           "../Data/Historial Whaling data/R/Soviet_SRW_catches.csv"))

# load files, put in spatial format and project
# filtering on months based on range across srw.data being 6 to 10. +/- 3 to 5
# potential range of months for whaling will therefore be january (1) to July (7)
# the soviet and yankee data are not under the same format so we are making two different pathways in the function
whaling_df <- tibble(files) %>% 
  mutate(source = c("Yankee", "Soviet")) %>% 
  mutate(data = map(files, read.csv)) %>% 
  dplyr::select(-files) %>% 
  mutate(data = map2(data, source, function(d, s){
    if(s == "Yankee"){
      d <- d %>% # select only the months we are interested in
        filter(Month %in% c(1:7), Lat <= -30) %>% 
        st_as_sf(coords = c("Long", "Lat"), crs = 4326) %>% # convert sf
        st_transform(crs = st_crs(proj_antarctic)) %>% # project
        mutate(decade = floor(Year/10)*10) %>% 
        dplyr::select(decade, Month, Year, geometry) %>% 
        dplyr::rename(month = Month, year = Year)
      return(d)
    }
    if(s == "Soviet"){
      d <- d %>% # select only the months we are interested in
        filter(month %in% c(1:7), Latitude <= -30) %>% 
        st_as_sf(coords = c("Longitude","Latitude"), crs = 4326) %>% 
        st_transform(crs = st_crs(proj_antarctic)) %>% 
        mutate(year = year + 1900,
               decade = floor(year/10)*10) %>% 
        dplyr::select(decade, month, year, geometry)
      return(d)
    }
  })) %>% 
  unnest(data)

# number of records (each row is one whale)
nrow(whaling_df)
# temporal distribution of records
summary(whaling_df$year)
# number of records by source
table(whaling_df$source)
# number of records by source and by decade
whaling_df %>% 
  group_by(source, decade) %>% 
  dplyr::summarize(nb_records = n())

save(whaling_df, file = "./Outputs/whaling_df.RData")
```

Selecting only the points that fall within the feeding bubbles
```{r}
# in this dataframe, the whaling points overlapping with multiple feeding bubbles will be repeated
whaling_inbubbles_df <- whaling_df %>% 
  st_as_sf() %>% 
  mutate(index = seq(1:nrow(whaling_df))) %>% 
  st_intersection(srw.buf) %>% 
  mutate(Locality = sub(" ", ".", Locality1))

# number of records left
length(unique(whaling_inbubbles_df$index))

save(whaling_inbubbles_df, file = "./Outputs/whaling_inbubbles_df.RData")
```

## Map whaling catches over feeding areas

```{r, fig.cap = "Population level feeding areas overlapped with whaling data", fig.align="center", fig.height = 8, fig.width = 8, warning = F}
g <- ggplot() +
  coord_sf(label_graticule = "") +
  theme_light() +
  theme(legend.position = "") +
  xlab("") + ylab("") +
  geom_tile(data = breeding.bin, aes(x, y, fill = Locality), alpha = 0.7) +
  geom_sf(data = world, col = "transparent", fill = "grey15") +
  facet_wrap(~Locality, ncol = 3) +
  geom_sf(data = whaling_df, aes(geometry = geometry, col = source), size = 0.5, alpha = 0.8) +
  scale_color_manual(values = c("grey", "black"))

ggsave(g, file = "./Outputs/Population_feeding_grounds_whaling.png", width = 8, height = 8)
g
```

```{r, fig.cap = "Population level feeding areas overlapped with whaling data by area", fig.align="center", fig.height = 8, fig.width = 8, warning = F}
g <- ggplot() +
  coord_sf(label_graticule = "") +
  theme_light() +
  theme(legend.position = "") +
  xlab("") + ylab("") +
  geom_tile(data = breeding.bin, aes(x, y, fill = Locality), alpha = 0.7) +
  geom_sf(data = world, col = "transparent", fill = "grey15") +
  facet_wrap(~Locality, ncol = 3) +
  geom_sf(data = whaling_inbubbles_df, aes(geometry = geometry), size = 0.5, alpha = 0.8)

ggsave(g, file = "./Outputs/Population_feeding_grounds_whaling_inzones.png", width = 8, height = 8)
g
```

## Calculate overlap

```{r}
# grid example used to rasterize points and calculate overlap
breeding.ras

# convert feeding areas 50 % and 25 % to polygons
breeding.pol <- breeding.bin %>% 
  pivot_longer(cols = c(feeding0.25, feeding0.50), names_to = "feeding_ground", values_to = "feeding_value") %>% 
  filter(!is.na(feeding_value)) %>% 
  dplyr::select(-feeding_value) %>% 
  group_by(Locality, feeding_ground) %>% 
  nest() %>% 
  mutate(feeding_pol = map(data, function(d){
    d %>% 
      st_as_sf(coords = c("x", "y"), crs = st_crs(proj_antarctic)) %>% 
      rasterize(., breeding.ras) %>% 
      st_as_stars() %>% 
      st_as_sf(as_points = FALSE, merge = TRUE) %>% 
      st_as_sf()
  })) %>% 
  unnest(feeding_pol) %>% 
  st_as_sf()
st_agr(breeding.pol) <- "constant"

# calculate intersection
whaling_overlap_df <- whaling_inbubbles_df %>% 
  dplyr::select(source, decade, month, year, index, Locality) %>% 
  group_by(Locality) %>% 
  nest() %>% 
  mutate(overlap = map2(Locality, data, function(l, d){
            st_agr(d) <- "constant"
            d %>% 
              st_intersection(breeding.pol[breeding.pol$Locality == l, ])}),
        percent_overlap = map2(data, overlap, function(all_p, in_p){
            in_p %>% 
              group_by(feeding_ground) %>% 
              dplyr::summarize(nb_overlap_points = n(), # nb of overlap cells by locality then divide by total number of whaling points for that locality
                percent_overlap = round(nb_overlap_points / nrow(all_p) *100, 0))})) 

# set order of locality in which we want them to appear
locs <- c("Argentina", "Argentina_d15Nlow", "Argentina_d15Nhigh", "Brazil", "South.Africa", "Southwest.Australia", "Southeast.Australia", "NZ.Mainland", "Auckland.Islands")

# convert to csv to export
whaling_overlap_df %>% 
  unnest(percent_overlap) %>% 
  dplyr::select(-c(data, overlap, nb_overlap_points, geometry)) %>% 
  mutate(Locality = factor(Locality, levels = locs)) %>%
  arrange(Locality) %>% 
  pivot_wider(names_from = feeding_ground, values_from = percent_overlap) %>%
  replace(is.na(.), 0) %>% 
  print(n = Inf) %>% 
  write.csv(file = "./Outputs/Table_overlap_whalingrecords.csv", row.names = F)

```

## Calculate correlation nb_records ~ proba feeding

Summing the number of whaling records per pixel and then looking at the correlation between nb of records and mean probability of feeding
```{r}
whaling_gridded <- ddply(whaling_inbubbles_df, ~Locality, function(d){
  # rasterize by summing counts of animals within cells
  # each record is one whale so add a column to indicate that number and then sum
  d$count <- 1
  # set the grid to the relevant breeding ground
  grid <- breeding.ras[[d$Locality[1]]]
  # rasterize
  d <- d %>% 
      st_as_sf() %>% 
      rasterize(., grid, field = "count", fun = "sum") 
  # combine the probabilities from grid to the rasterized whaling records from d
  out <- cbind(as.data.frame(grid), as.data.frame(d))
  names(out) <- c("proba_feeding", "nb_whaling")
  out <- out %>% 
    filter(!is.na(proba_feeding)) %>% # remove cells outside feeding bubbles
    mutate_all(~replace(., is.na(.), 0)) # replace na by zeros in nb_whaling because it means there was no records in that cell and we want to account for that
  return(out)
})

save(whaling_gridded, file = "./Outputs/whaling_gridded.RData")
```


```{r}
ggplot(whaling_gridded, aes(nb_whaling, proba_feeding, col = Locality)) +
  geom_point(alpha = 0.8) +
  facet_wrap(~Locality, scales = "free")

# log-transformed version
g <- ggplot(whaling_gridded, aes(log(nb_whaling + 1), log(proba_feeding))) +
  geom_point(alpha = 0.8) +
  facet_wrap(~Locality, scales = "free") +
  stat_smooth(method='lm', formula = y~poly(x,2), col= "grey20")
g

ggsave(g, file = "./Outputs/Correlation_whaling_proba.png", width = 8, height = 8)

# calculate spearman coefficient of correlation (data not gaussian)
cor_fun <- function(df) cor.test(df$proba_feeding, df$nb_whaling, method="spearman", exact = F) %>% broom::tidy()

Fun_whalingcor <- function(df) nest(df) %>% 
  mutate(model = purrr::map(data, cor_fun)) %>% 
  dplyr::select(-data) %>% 
  unnest(cols = c(model))

whaling_gridded %>% 
  group_by(Locality) %>% 
  Fun_whalingcor() 
```


Save data
```{r}
save(srw.data, file = "./Outputs/srw.data.RData")
```

















